Analyzing the Effect of Surface in Tennis Grand Slams

### 2018-09-14 contest/graphics/graphics.Rmd
add spainiards clay and update graphics

grid.arrange(p1,p2, p3, nrow=1, widths = c(.29, .29, .42),
             top = textGrob("Home court advantage debunked: win proportion by country",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left"))

@@

```{r}
filter(gs_players, round == "R128", ioc == "ESP") %>% 
  group_by(tournament) %>% 
  summarise(avg_ranking = mean(rank))

spain_ranks <- filter(gs_players, ioc == "ESP") %>% 
  group_by(tournament, round) %>% 
  summarise(avg_ranking = mean(rank))

ggplot(spain_ranks) + 
  geom_line(aes(round, avg_ranking, color=tournament, group=tournament)) + 
  labs(title = "Average ranking among Spaniards")

@@

## Spain proportion of wins across the tournaments

```{r}
spain_wins <- filter(gs_players, ioc == "ESP") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

ggplot(spain_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line(cex=2) + 
  geom_point() +
  labs(title = "Proportion of wins among Spaniards") +
  geom_errorbar(aes(ymin = win_prop - 2*win_prop_se,
                    ymax = win_prop + 2*win_prop_se),
                width=.2,
                 position=position_dodge(0.05))

>>>>

grid.arrange(p1,p2, p3, nrow=1, widths = c(.29, .29, .42),
             top = textGrob("Home court advantage debunked: win proportion by country",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left",
                           gp=gpar(fontsize=20, fontfamily="serif")))

@@

It is also commonly thought that Spaniards play better on clay. We are intersted in whether Spaniards win the French Open more than they win in other tournaments. It does appear that Spaniards are winning the French Open more than they are winning other tournaments. But, this result is not significant. In addition, the ranking of Spaniards in the French Open is, on average, higher than other tournmanents, which may help explain this common misconseption, although this result is not significant either.

spain_wins <- filter(gs_players, ioc == "ESP", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()),
            avg_ranking = mean(rank),
            rank_se = sd(rank)/sqrt(n()))

p1 <- ggplot(spain_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  labs(title = "Win proportion among Spaniards",
       # caption = "Note: we only look at the first 4 rounds due to decreasing sample size",
       x = "", y = "Win proportion") + 
  scale_color_manual(guide=FALSE, values=tournament_colors) + my_theme  
  # geom_point() + 
  # labs(title = "Proportion of wins among Spaniards") +
  # geom_errorbar(aes(ymin = win_prop - win_prop_se,
  #                   ymax = win_prop + win_prop_se),
  #               width=.2,
  #                position=position_dodge(0.05))

p2 <- ggplot(spain_wins, aes(round, avg_ranking, color=tournament, group=tournament)) + 
  geom_line() + 
  labs(title = "Average ranking among Spaniards",
       x = " ",
       y = "Average rank") + 
  scale_color_manual(" ", values=tournament_colors) + my_theme 
# +
#   geom_errorbar(aes(ymin = avg_ranking - 1.96*rank_se,
#                     ymax = avg_ranking + 1.96*rank_se),
#                 width=.2,
#                  position=position_dodge(0.05))

grid.arrange(p1,p2, nrow=1, widths = c(.44, .56),
             top = textGrob("Spaniards perform better on Clay: debunked",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left",
                           gp=gpar(fontfamily="serif")))

@@

### 2018-09-14 contest/surface-type/paper-tennis.Rmd
add spainiards clay and update graphics

>>>>

## Spaniards on clay

It is also commonly thought that Spaniards play better on clay. We are intersted in whether Spaniards win the French Open more than they win in other tournaments. It does appear that Spaniards are winning the French Open more than they are winning other tournaments. But, this result is not significant. In addition, the ranking of Spaniards in the French Open is, on average, higher than other tournmanents, which may help explain this common misconseption, although this result is not significant either.

spain_wins <- filter(gs_players, ioc == "ESP", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()),
            avg_ranking = mean(rank),
            rank_se = sd(rank)/sqrt(n()))

p1 <- ggplot(spain_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  labs(title = "Win proportion among Spaniards",
       # caption = "Note: we only look at the first 4 rounds due to decreasing sample size",
       x = "", y = "Win proportion") + 
  scale_color_manual(guide=FALSE, values=tournament_colors) + my_theme  
  # geom_point() + 
  # labs(title = "Proportion of wins among Spaniards") +
  # geom_errorbar(aes(ymin = win_prop - win_prop_se,
  #                   ymax = win_prop + win_prop_se),
  #               width=.2,
  #                position=position_dodge(0.05))

p2 <- ggplot(spain_wins, aes(round, avg_ranking, color=tournament, group=tournament)) + 
  geom_line() + 
  labs(title = "Average ranking among Spaniards",
       x = " ",
       y = "Average rank") + 
  scale_color_manual(" ", values=tournament_colors) + my_theme 
# +
#   geom_errorbar(aes(ymin = avg_ranking - 1.96*rank_se,
#                     ymax = avg_ranking + 1.96*rank_se),
#                 width=.2,
#                  position=position_dodge(0.05))

grid.arrange(p1,p2, nrow=1, widths = c(.44, .56),
             top = textGrob("Spaniards perform better on Clay: debunked",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left",
                           gp=gpar(fontfamily="serif")))
